\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-11-25 tex/gitlinks.tex
tex updates

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/b677d4b247da6117fd73f66b451765dac961c376/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

>>>>

\newcommand{\codelink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/figures/#1.py}{\codeicon}\,\,}
\newcommand{\animlink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/figures/#1.gif}{\animicon}\,\,}
\newcommand{\prooflink}[1]{\href{https://github.com/rodluger/starry_process/blob/09b20fbf11c7967ce0510c149223cc7406ca9eca/tex/tests/#1.py}{\raisebox{-0.1em}{\input{tests/#1.tex}}}}
\newcommand{\cilink}[1]{\href{https://dev.azure.com/rodluger/starry_process/_build}{#1}}

### 2020-11-25 tex/ms.tex
tex updates

\section{Spherical Harmonics}
\label{sec:ylm}
%
The spherical harmonics are a complete set of orthogonal functions on the
surface of the sphere and, as such, are a convenient basis in which to construct
our gaussian process. As in \citet{Luger2019}, here we describe the intensity
on the surface of a star as an expansion in the real spherical harmonics, which is
fully specified by a set of spherical harmonic coefficients
$\{ y^l_m \}$, where $l \in [0, l_{\mathrm{max}}]$ is known as
the degree and
$m \in [-l, l]$ is known as the order, with $l_{\mathrm{max}}$ equal to the highest
degree in the expansion.
It is often convenient to collect the spherical harmonic coefficients of
a given expansion into a vector
%
\begin{align}
    \label{eq:y}
    \mathbf{y}
     & =
    \left(
    \mathbf{y}^0 \,\,\,
    \mathbf{y}^1 \,\,\,
    \cdots \,\,\,
    \mathbf{y}^{l_\mathrm{max}}
    \right)^\top
    \nonumber \\
     & =
    \left(
    y^0_0 \,\,\,
    \,\,\,
    y^1_{-1} \,\,\,
    y^1_{0} \,\,\,
    y^1_{1} \,\,\,
    \,\,\,
    \cdots \,\,\,
    \,\,\,
    y^{l_\mathrm{max}}_{-l_\mathrm{max}} \,\,\,
    \cdots \,\,\,
    y^{l_\mathrm{max}}_{l_\mathrm{max}} \,\,\,
    \right)^\top
    \quad.
\end{align}
%
\citet{Luger2019} derived the expression for the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe from a star rotating about a fixed axis
if the stellar surface intensity is well described by the spherical harmonic
coefficient vector $\mathbf{y}$:
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \pmb{\mathcal{A}} \, \mathbf{y}
    \quad,
\end{align}
%
where $\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is implicitly
a function of the stellar inclination $i$, the stellar
rotation period $P$, the stellar limb darkening coefficients, etc.
For reference, the rows of $\pmb{\mathcal{A}}$ are given by
Equation~(18) in \citet{Luger2019}.

An in-depth treatment of the spherical harmonics is beyond the scope of this
paper. For more information, including closed-form expressions for
the spherical harmonics
and details on their usage in modeling stellar surfaces, the reader is
encouraged to consult \citet{Luger2019}.

\section{A Gaussian Process for Star Spots}
\label{sec:gp}
%

@@

Let
$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_K \right)^\top$
denote a vector of $K$ flux measurements at times
$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_K \right)^\top$.
Conditioned on certain physical properties of the
star
(such as its inclination, rotation period, etc.), $\pmb{\theta}_\star$,
and on certain properties of the  star spots
(such as their sizes, positions, etc.), $\pmb{\theta}_\bullet$,
we wish to compute the mean $\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ and
covariance $\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$
of $\mathbf{f}$, which together fully specify our GP in flux.
%
As with any random variable, the mean and covariance may be computed from
the expectation value of $\mathbf{f}$ and
$\mathbf{f}\,\mathbf{f}^\top$, respectively:
%
\begin{align}
    \label{eq:mean}
    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big]
    \\
    \label{eq:cov}
    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
    \quad.
\end{align}
%
Given the linear relationship between flux and spherical harmonic
coefficients (Equation~\ref{eq:fAy}),
we may write the mean and covariance of our flux GP as
%
\begin{align}
    \label{eq:mean_f}
    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
    \\
    \label{eq:cov_f}
    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta}_\star)
    \quad,
\end{align}
%
where
%
\begin{align}
    \label{eq:mean_y}
    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    \\
    \label{eq:cov_y}
    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta}_\bullet)
\end{align}
%
are the mean and covariance of the GP in the spherical harmonics basis.
The bulk of the math in this paper is devoted to computing
the expectations in the expressions above, which
are given by the integrals
%
\begin{align}
    \label{eq:exp_y}
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \\
    \label{eq:exp_yy}
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \quad,
\end{align}
%
where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
distribution of features on the surface  (i.e., the size and location of star spots)
and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
function (PDF).

In the sections that follow we will show that for suitable choices of $\mathbf{y}(\mathbf{x})$

@@

As we are specifically interested in modeling the effect of star spots
on stellar light curves, we let

@@

    \mathbf{x} =

@@

    c_0 \,\, \cdots \,\, c_{N-1} \,\,
    ,
    \lambda_0 \,\, \cdots \,\, \lambda_{N-1} \,\,
    ,
    \phi_0 \,\, \cdots \,\, \phi_{N-1} \,\,
    ,
    r_0 \,\, \cdots \,\, r_{N-1} \,\,

@@

and
    \label{eq:RRs}
    \mathbf{y}(\mathbf{x}) =
    \sum_{n=0}^{N-1}
    c_n
    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)
    \mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)
    \mathbf{s}(r_n)

@@

where $c_n$ is the contrast of the $n^\mathrm{th}$ spot,
$\lambda_n$ is its longitude, $\phi_n$ is its latitude,
$r_n$ is its radius, and $N$ is the total number of spots.
The vector function $\mathbf{s}(r_n)$
returns the spherical harmonic expansion of a negative unit brightness
circular spot of radius $r_n$ at $\lambda = \phi = 0$,
$\mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)$ is the Wigner matrix that rotates the
expansion about $\hat{\mathbf{x}}$ such that the spot is centered at a
latitude $\phi_n$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)$ is the Wigner
matrix that then rotates the
expansion about $\hat{\mathbf{y}}$ such that the spot is centered at a
longitude $\lambda_n$; these three functions are detailed in the sections below.

@@

Equation~(\ref{eq:RRs}) thus provides a way of converting a random variable
$\mathbf{x}$ describing the size, brightness, and position of spots to the
corresponding representation in terms of spherical harmonics.

@@

Note, importantly, that we are not interested in any specific value of
$\mathbf{y}$; rather, we would like to know its expectation value under
the probability distribution governing the different spot properties $\mathbf{x}$,
i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$.

@@

For simplicity, we assume that $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$
is separable in each of the four spot properties, and that all of the spots

@@

    p(c_n \, \big| \, \pmb{\theta}_{c}) \,

@@

    p(r_n \, \big| \, \pmb{\theta}_{r})

@@

    \pmb{\theta}_{c} \, \,

@@

is the vector of yet-to-be-specified hyperparameters describing the process.

@@

This allows us to rewrite the expectation integrals (\ref{eq:exp_y})
and (\ref{eq:exp_yy}) as

@@

    N \mathbf{e}_c

@@

    N \mathbf{E}_c

@@

    \mathbf{e}_r

@@

    \mathbf{s}(r) \,
    p(r \, \big| \, \pmb{\theta}_{r}) \,
    \mathrm{d}r

@@

@@

    \mathbf{e}_c

@@

    c \,

@@

    p(c \, \big| \, \pmb{\theta}_{c}) \,
    \mathrm{d}c

@@

    \mathbf{E}_r

@@

@@

    \mathbf{E}_c

@@

    c^2 \,

@@

    p(c \, \big| \, \pmb{\theta}_c)
    \mathrm{d}c

@@

We devote the next section to the computation of these eight
integrals.

\section{The Expectation Integrals}
\label{sec:integrals}
\subsection{The Size Integral}
\label{sec:size}+
%
Our goal in this section is to compute the first and second moments
of the size distribution ($\mathbf{e}_\phi$, $\mathbf{E}_\phi$) under
a suitable spherical harmonic expansion $\mathbf{s}(r)$ of the spot profile
and a suitable probability distribution function for the spot radius,
$p(r \, \big| \, \pmb{\theta}_{r})$.

\subsubsection{Spot profile}
%
We model the brightness $b$ an angle $\vartheta$ away from the
center of a spot of negative unit intensity and radius $r$ be
%
\begin{align}
    \label{eq:brvartheta}
    b(r; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{r-\vartheta}{s}\right)} - 1
\end{align}
%
for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
inverted top-hat function with half-width equal to $r$,
corresponding to a circular spot of uniform intensity. For $s > 0$, each half
of $b$ is a sigmoid with half-width at half-minimum equal to $r$.
In our implementation of the algorithm we choose $s = 0.2^\circ$,
which is small compared to features of interest but not too small as to
create numerical issues when computing model gradients (which would be
undefined at the spot boundary if the spot profile were truly an inverted top-hat).

>>>>

\section{A Gaussian Process for Star Spots}

\subsection{Spherical Harmonics}
\label{sec:ylm}
%
The spherical harmonics are a complete set of orthogonal functions on the
surface of the sphere and, as such, are a convenient basis in which to construct
our gaussian process. As in \citet{Luger2019}, here we describe the intensity
on the surface of a star as an expansion in the real spherical harmonics.
%
\citet{Luger2019} derived the expression for the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe from a star rotating about a fixed axis
if the stellar surface intensity is well described by the spherical harmonic
coefficient vector $\mathbf{y}$:
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \pmb{\mathcal{A}} \, \mathbf{y}
    \quad,
\end{align}
%
where $\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is implicitly
a function of the stellar inclination $i$, the stellar
rotation period $P$, the stellar limb darkening coefficients, etc.
(see Appendix~\ref{sec:starry} for details).

\subsection{Computing the GP}
\label{sec:gp}
%

@@

Let
$\mathbf{f} = \left( f_0 \, f_1 \, \cdots \,  f_{K-1} \right)^\top$
denote a vector of $K$ flux measurements at times
$\mathbf{t} = \left( t_0 \,  t_1 \,  \cdots \, t_{K-1} \right)^\top$.
Conditioned on certain physical properties of the
star
(such as its inclination, rotation period, etc.), $\pmb{\theta}_\star$,
and on certain properties of the  star spots
(such as their sizes, positions, etc.), $\pmb{\theta}_\bullet$,
we wish to compute the mean $\pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$ and
covariance $\pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)$
of $\mathbf{f}$, which together fully specify our GP in flux.
As with any random variable, the mean and covariance may be computed from
the expectation values of $\mathbf{f}$ and
$\mathbf{f}\,\mathbf{f}^\top$, respectively:
    \label{eq:mean}
    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big]
    \label{eq:cov}
    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \pmb{\theta}_\star, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet) \pmb{\mu}^\top(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
Given the linear relationship between flux and spherical harmonic
coefficients (Equation~\ref{eq:fAy}),
we may write the mean and covariance of our flux GP as
    \label{eq:mean_f}
    \pmb{\mu}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
    \label{eq:cov_f}
    \pmb{\Sigma}(\pmb{\theta}_\star, \pmb{\theta}_\bullet)
     & = \pmb{\mathcal{A}}(\pmb{\theta}_\star) \, \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet) \, \pmb{\mathcal{A}}^\top(\pmb{\theta}_\star)
    \label{eq:mean_y}
    \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    \label{eq:cov_y}
    \pmb{\Sigma}_{\mathbf{y}}(\pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] - \pmb{\mu}_{\mathbf{y}}(\pmb{\theta}_\bullet) \pmb{\mu}_{\mathbf{y}}^\top(\pmb{\theta}_\bullet)
are the mean and covariance of the GP in the spherical harmonics basis.
The bulk of the math in this paper (Appendix~\ref{sec:integrals})
is devoted to computing
the expectations in the expressions above, which
are given by the integrals
    \label{eq:exp_y}
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \label{eq:exp_yy}
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
where $\mathbf{x}$ is a random vector-valued variable corresponding to a particular
distribution of features on the surface
and $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability density
function (PDF).
In the Appendix we show that for suitable choices of $\pmb{\theta}_\bullet$,
$\mathbf{y}(\mathbf{x})$,

@@

In particular, we let the spot hyperparameters of our GP be

@@

    \pmb{\theta}_\bullet

@@

    N
    c
    \alpha
    \beta
    r
    \Delta

@@

@@

where $N$ is the number of star spots, $c$ is their contrast,
$\alpha$ and $\beta$ describe the latitudinal
distribution of the spots, and $r$ and $\Delta$ are the mean and
half-width, respectively, of the distribution of spot radii.

@@

The PDF for the radius is a uniform distribution
between $r - \Delta$ and $r + \Delta$ (Appendix \ref{sec:size}),
while
the PDF for the latitude $\phi$ of the spots is a Beta distribution in
$\cos\phi$ with shape parameters $\alpha$ and $\beta$ (Appendix \ref{sec:lat-lon}).
The spot longitude is assumed to be uniformly distributed. Finally,
the PDFs for the spot contrast and the number of spots are delta functions
centered at $c$ and $N$, respectively.

\xxx{More soon.}

\subsection{Marginalizing over inclination}
%
The equations for the mean and covariance of the flux GP
(Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
on specific values of the stellar properties $\pmb{\theta}_\star$ and the
spot properties $\pmb{\theta}_\bullet$. To obtain the posterior distribution
for these parameters, we must typically resort to numerical sampling techniques,
which often scale steeply with the number of parameters. It is therefore generally
desirable to keep the total number of parameters small, especially when
employing the GP in an ensemble setting.
In such a setting, we might have light curves from $N_\star$ stars, all of which
we believe to have similar spot properties (perhaps because they have
similar spectral types and rotation periods, for example).
The total number of parameters in our problem is therefore
%
\begin{align}
    N_\theta = N_\star \, \mathrm{dim}(\pmb{\theta}_\star) + 5
    \quad,
\end{align}
%
since each of the stars will have their own set of stellar properties
$\pmb{\theta}_\star$ but will all share the same 5 spot properties
$\pmb{\theta}_\bullet$ (by assumption).
In a typical problem, the stellar parameters might be
%
\begin{align}
    \pmb{\theta}_\star =
    \left(
    i \,\,\,
    P \,\,\,
    u_1 \,\,\,
    u_2
    \right)^\top
    \quad,
\end{align}
%
i.e., the inclination, the rotational period, and some (usually two) limb
darkening coefficients. Thus, for a reasonably sized ensemble of $N_\star=100$
stars, we would have to sample over $N_\theta = 405$ parameters.
While large, this number is certainly not absurd, especially by modern standards.
However, it does pose
a problem when considering how complex the posterior distribution for the
spot mapping problem can be. In addition to strong nonlinear degeneracies
between some of the parameters (such as the contrast $c$ and the
number of spots $N$), the posterior is often multimodal, especially in the
stellar inclinations. While modern sampling schemes such as Hamiltonian
Monte Carlo and Nested Sampling may in principle be able to deal with these
issues, in practice it can be very difficult to obtain convergence in a
reasonable amount of time.

Fortunately, there is a better way to reduce the number of parameters in
the problem: we can explicitly marginalize over the stellar inclination.
That is, we can write the first two moments of our flux GP as
%
\begin{align}
    \label{eq:mu_marg}
    \mathrm{E} \Big[ \mathbf{f} \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
     & =
    \int
    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    p(i) \, \mathrm{d}i \\
    \label{eq:cov_marg}
    \mathrm{E} \Big[ \mathbf{f} \, \mathbf{f}^\top \, \Big| \, \bar{\pmb{\theta}}_\star, \pmb{\theta}_\bullet \Big]
     & =
    \int
    \pmb{\mathcal{A}}(i, \bar{\pmb{\theta}}_\star) \,
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
    \pmb{\mathcal{A}}^\top(i, \bar{\pmb{\theta}}_\star)
    p(i) \, \mathrm{d}i    \quad,
\end{align}
%
where the expectations inside the integrals are given by
Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively, and
$\bar{\pmb{\theta}}_\star$ are all 0524the stellar parameters \emph{except}
the inclination, i.e.,
%
\begin{align}
    \bar{\pmb{\theta}}_\star = \left(
    P \,\,\,
    u_1 \,\,\,
    u_2
    \right)^\top
    \quad,
\end{align}
%
where perhaps $P$ is fixed for each star and the limb darkening coefficients
are shared among all stars, as described above.
%
Thus, if we are able to perform the integrals in Equations~(\ref{eq:mu_marg})
and (\ref{eq:cov_marg}), we can \emph{dramatically} reduce the number of
parameters in our ensemble problem.
%
As we show in Appendix~\ref{sec:inclination}, if we assume that stellar
inclinations are distributed isotropically, these integrals
do in fact have closed-form solutions,
since they are (again) integrals over Wigner matrices.
%

\section{Stuff}

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/nullspace.pdf}
        \oscaption{nullspace}{%
            The null space.
            \label{fig:nullspace}
        }
    \end{centering}
\end{figure}

%
%
%
%

%
%
%
%

\section{Notation}
\label{sec:notation}
%
Unless otherwise noted, we adopt
the following conventions throughout this paper:
integers are represented by italic uppercase letters (i.e., $N$),
scalars are represented by italic lowercase
letters (i.e., $x$), column vectors are
represented by boldface lowercase letters
($\mathbf{x}$), and matrices and other linear operators are represented
by boldface capital letters ($\mathbf{X}$). In general, the elements of a vector
$\mathbf{x}$ are denoted $x_i$ and the elements of a matrix $\mathbf{X}$
are denoted $X_{i,j}$. The exception to this is if the vector or matrix
represents a quantity in the spherical harmonic basis, in which case we use
\emph{two} indices to represent a vector element, $x^l_m$, and \emph{four}
indices to represent a matrix element, $X^{l,l'}_{m,m'}$.
%
The upper indices corresponds to the spherical harmonic degree,
$l \in [0, l_{\mathrm{max}}]$ and $l' \in [0, l_{\mathrm{max}}]$,
while the lower indices correspond to the
spherical harmonic order, $m \in [-l, l]$ and $m' \in [-l', l']$.
%
Vector elements are arranged in order of increasing $l$ and,
within each $l$, in order of increasing $m$.
For example, a vector $\mathbf{x}$
representing a quantity in the spherical harmonic basis up to degree
$l_\mathrm{max}$ has components given by
%
\begin{align}
    \mathbf{x}
     & =
    \left(
    x^0_0 \,\,\,
    \,\,\,\,\,\,
    x^1_{-1} \,\,\,
    x^1_{0} \,\,\,
    x^1_{1} \,\,\,
    \,\,\,\,\,\,
    \cdots \,\,\,
    \,\,\,\,\,\,
    x^{l_\mathrm{max}}_{-l_\mathrm{max}}
    \cdots \,\,
    x^{l_\mathrm{max}}_{l_\mathrm{max}} \,\,\,
    \right)^\top
    \quad,
\end{align}
%
while a matrix $\mathbf{X}$ in the same basis has components given by
%
\begin{align}
    \setstackgap{L}{1.25\baselineskip}
    \fixTABwidth{T}
    \mathbf{X} =
    \parenMatrixstack{
    X^{0,0}_{0,0}  &  & X^{0,1}_{0,-1}  & X^{0,1}_{0,0}  & X^{0,1}_{0,1}  &        \\
                   &  &                 &                &                &        \\
    X^{1,0}_{-1,0} &  & X^{1,1}_{-1,-1} & X^{1,1}_{-1,0} & X^{1,1}_{-1,1} &        \\
    X^{1,0}_{0,0}  &  & X^{1,1}_{0,-1}  & X^{1,1}_{0,0}  & X^{1,1}_{0,1}  &        \\
    X^{1,0}_{1,0}  &  & X^{1,1}_{1,-1}  & X^{1,1}_{1,0}  & X^{1,1}_{1,1}  &        \\
                   &  &                 &                &                & \ddots
    }\quad.
\end{align}
%
For completeness, the element of a spherical harmonic vector $\mathbf{x}$ with
degree $l$ and order $m$ is at (flattened) index
%
\begin{align}
    \label{eq:n}
    n = l^2 + l + m
    \quad.
\end{align}
%
Conversely, the element at (flattened) index $n$ has degree and order
%
\begin{align}
    \label{eq:lm}
    \begin{split}
        l & = \floor{\sqrt{n}}
        \\
        m & = n - l^2 - l
        \quad,
    \end{split}
\end{align}
%
respectively.
%
Note, finally, that our use of upper and lower indices is purely a
notational convenience,
and should not be confused with
exponentiation or a distinction between covariant and contravariant
tensors. It should also not be confused with the notation used for the complex
spherical harmonics, which also uses upper and lower indexing.
%

\section{Computing the flux}
\label{sec:starry}
\xxx{Coming soon.}

\section{The Expectation Integrals}
\label{sec:integrals}

Our goal in this section is to find closed-form solutions to the
first and second moments of the spherical harmonic representation of the
stellar surface $\mathbf{y}$,
%
\begin{align}
    \label{eq:exp_y_app}
    \mathrm{E} \Big[ \mathbf{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \\
    \label{eq:exp_yy_app}
    \mathrm{E} \Big[ \mathbf{y} \, \mathbf{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big]
     & =
    \int \mathbf{y}(\mathbf{x} ) \mathbf{y}^\top(\mathbf{x} ) \, p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)\mathrm{d}\mathbf{x}
    \quad,
\end{align}
%
which are linearly related to the mean and the covariance of our GP in flux
(\S\ref{sec:gp}).
%
Recall that $\mathbf{x}$ is a vector of parameters describing the particular
configuration of features on the surface of a star, and
$p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$ is its probability
density function conditioned on hyperparameters $\pmb{\theta}_\bullet$,
which describe the \emph{distribution} of the features on the surface
of one or many stars.
%
As we are specifically interested in modeling the effect of star spots
on stellar light curves, we let
%
\begin{align}
    \mathbf{x} =
    \left(
    \eta \,\,\,\,
    \xi_0 \,\, \cdots \,\, \xi_{\eta-1} \,\,\,\,
    \lambda_0 \,\, \cdots \,\, \lambda_{\eta-1} \,\,\,\,
    \phi_0 \,\, \cdots \,\, \phi_{\eta-1} \,\,\,\,
    \rho_0 \,\, \cdots \,\, \rho_{\eta-1}
    \right)^\top
\end{align}
%
and
%
\begin{align}
    \label{eq:RRs}
    \mathbf{y}(\mathbf{x}) =
    \sum_{n=0}^{\eta-1}
    \xi_n
    \,
    \mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)
    \,
    \mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)
    \,
    \mathbf{s}(\rho_n)
    \quad,
\end{align}
%
where $\eta$ is the total number of spots,
$\xi_n$ is the contrast of the $n^\mathrm{th}$ spot,
$\lambda_n$ is its longitude, $\phi_n$ is its latitude,
and $\rho_n$ is its radius.
The vector function $\mathbf{s}(r_n)$
returns the spherical harmonic expansion of a negative unit brightness
circular spot of radius $r_n$ at $\lambda = \phi = 0$,
$\mathbf{R}_{\hat{\mathbf{x}}}(\phi_n)$ is the Wigner matrix that rotates the
expansion about $\hat{\mathbf{x}}$ such that the spot is centered at a
latitude $\phi_n$, and $\mathbf{R}_{\hat{\mathbf{y}}}(\lambda_n)$ is the Wigner
matrix that then rotates the
expansion about $\hat{\mathbf{y}}$ such that the spot is centered at a
longitude $\lambda_n$; these three functions are detailed in the sections below.
%
Equation~(\ref{eq:RRs}) thus provides a way of converting a random variable
$\mathbf{x}$ describing the size, brightness, and position of spots to the
corresponding representation in terms of spherical harmonics.
%
Note, importantly, that we are not interested in any specific value of
$\mathbf{y}$; rather, we would like to know its expectation value under
the probability distribution governing the different spot properties $\mathbf{x}$,
i.e., $p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$.
%

For simplicity, we assume that
the total number of spots is fixed to a value $N$, i.e.,

@@

    p(\eta \, \big| \, \pmb{\theta}_\bullet)
    \delta(\eta - N)

@@

where $\delta$ is the delta function.%
\footnote{%
    When modeling a single star using the GP, this assumption is justified by
    definition. It is less justified when the GP is used to model an ensemble of
    stars, where each star may have a different total number of spots $\eta$.
    However, as we argue in the text, $\eta$ is extremely difficult to constrain
    from light curves, in particular because of how degenerate it is with
    the spot contrast. In practice, we find that assuming that all stars in the
    ensemble have the same number of spots $N$ leads to higher variance in the
    estimate of $N$, but it does not lead to noticeable bias in $N$ or in any of
    the other hyperparameters.
We further asume that
$p(\mathbf{x} \, \big| \, \pmb{\theta}_\bullet)$
is separable in each of the four other spot properties, and that all of the spots

@@

    p(\xi_n \, \big| \, \pmb{\theta}_{\xi}) \,

@@

    p(\rho_n \, \big| \, \pmb{\theta}_{\rho})

@@

    N \, \,
    \pmb{\theta}_{\xi} \, \,

@@

is the vector of hyperparameters describing the process and
$\pmb{\theta}_{\xi}$,
$\pmb{\theta}_{\lambda}$,
$\pmb{\theta}_{\phi}$, and
$\pmb{\theta}_{\rho}$ are yet to be specified.

@@

This allows us to rewrite the expectation integrals (\ref{eq:exp_y_app})
and (\ref{eq:exp_yy_app}) as

@@

    N \mathbf{e}_\xi

@@

    N \mathbf{E}_\xi

@@

    \mathbf{e}_\rho

@@

    \mathbf{s}(\rho) \,
    p(\rho \, \big| \, \pmb{\theta}_{\rho}) \,
    \mathrm{d}\rho

@@

    \mathbf{e}_\rho \,

@@

    \mathbf{e}_\xi

@@

    \xi \,

@@

    p(\xi \, \big| \, \pmb{\theta}_{\xi}) \,
    \mathrm{d}\xi

@@

    \mathbf{E}_\rho

@@

    \mathbf{E}_\rho \,

@@

    \mathbf{E}_\xi

@@

    \xi^2 \,

@@

    p(\xi \, \big| \, \pmb{\theta}_\xi)
    \mathrm{d}\xi

@@

We devote the remainder of this section to the computation of these eight
integrals.

\subsection{The Radius Integral}
\label{sec:size}
%
Below we compute the first and second moments
of the radius distribution ($\mathbf{e}_\rho$, $\mathbf{E}_\rho$) under
a suitable spherical harmonic expansion $\mathbf{s}(\rho)$ of the spot profile
and a suitable probability distribution function for the spot radius,
$p(\rho \, \big| \, \pmb{\theta}_{\rho})$.

\subsubsection{Spot profile}
%
We model the brightness $b$ an angle $\vartheta$ away from the
center of a spot of negative unit intensity and radius $\rho$ as
%
\begin{align}
    \label{eq:brvartheta}
    b(\rho; \vartheta) & = \frac{1}{1 + \exp\left(\dfrac{\rho-\vartheta}{s}\right)} - 1
\end{align}
%
for some (constant) shape parameter $s$. In the limit $s \rightarrow 0$, $b$ approaches an
inverted top-hat function with half-width equal to $\rho$,
corresponding to a circular spot of uniform intensity. For $s > 0$, each half
of $b$ is a sigmoid with half-width at half-minimum equal to $\rho$.
In our implementation of the algorithm we choose $s = 0.2^\circ$,
which is small compared to features of interest but not too small as to
create numerical issues when computing model gradients (which would be
undefined at the spot boundary if the spot profile were truly an inverted top-hat).
